---
title: "RMarkdownExample"
author: "Andrew Allyn"
date: "5/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This is a short RMarkdown document to try to demonstrate a workflow that:
1. Uses the power of Git and GitHub for version controlled collaborative coding
2. Uses the power of Box for file backup and sharing files between lab members and the research department
3. Uses RMarkdown to nicely blend code, text and figures to create temporary results files that can be easily reviewed by lab members

## Workflow steps
1. I created a new repo on GitHub.com called "LabVingettes"
2. I cloned the repository, copying the https link, and then opened up a new version controlled RStudio project *importantly when I did this I made sure to create the new project within the GitHub folder in my personal box folder.* Though, this could have just as easily been done within the lab folder on Box. 
3. On my desktop, I then created a few subfolders within the new GitHub project folder ("Code", "Data", "Temp Results")
4. I edited the .gitignore file to include the Data folder and the Temp Results folder
5. I wrote this document
6. I staged, committed and pushed this RMarkdown code to GitHub
7. I opened a new script to render this RMarkdown code ("ExampleCodPlot_Render.R"), allowing me to set the output directory for the rendered file (Temp Results) that is different from the folder that this RMarkdown file exists (Code)
8. I staged, commited and pushed the render code to GitHub

## Start data exploration
A few preliminary things to start. First, we are going to source some functions that are in the ExampleSharedFunctions and some data in the ExampleSharedData folder. Importantly, this folder is currenty in my personal directory **and for now, you will need to update the "user name" in the code chunk below**. The idea would be that these folders would exist in the Research folder (e.g., Box-Research-Data or Box-Research-Functions). Given that, only the very beginning of the file path should change depending on the operating system. I imagine this would need to be at the beginning of all codes? *Is there a better way to do this??*
```{r}
# Detect the operating system
os.use<- .Platform$OS.type

# Set path to shared folders
user.name<- "Andrew Allyn/" 
shared.path<- switch(os.use, 
                     "unix" = paste("~/Box/", user.name, sep = ""),
                     "windows" = paste("C:/Box/", user.name, sep = ""))

# Source functions -- this is a little clumsy with the output text, but works
shared.funcs<- list.files(paste(shared.path, "ExampleSharedFunctions", sep = ""), full.names = TRUE)
sapply(shared.funcs, source)

# Load libraries
libraries.needed<- c("tidyverse", "here", "sf", "viridis", "SDMTools")
library_check(libraries.needed)
```

Now, let's see about reading in some example data (NOAA NEFSC bottom trawl survey data for Altantic cod), which is stored within the project data folder. Maybe using the here library? And then reading in a shapefile, which would be located in some shared folder.
```{r}
# Reading in cod data -- would want to try this with someone else's copy of this folder?
cod.dat<- read_csv(here("Data", "NOAANEFSCBottomTrawl_CodObs.csv"))
summary(cod.dat)

# Let's also read in the NELME shapefile, which is located in a different folder
# Option one...
nelme<- st_read(paste(shared.path, "ExampleSharedData/Shapefiles/NELME_sf.shp", sep = ""))
ggplot() +
  geom_sf(data = st_simplify(nelme, dTolerance = 0.001))

# That worked -- I don't think we can use the "here" piece because "here" is going to be looking inside this RProject and its corresponding folders. The ExampleSharedData folder, however, will be located outside this RProject.
```

Alright, so seemingly we are able to read in some data that is within the RProject Data folder and we are also able to read in a shared shapefile of the Northeast Large Marine Ecosystem. Next, let's make a quick plot of Cod Center of Gravity and save the figure to the Temp Results folder in this project...
```{r}
# The center of gravity will be calculated for each year and season independently. So, looking at a replicated process and either loops OR using new tidyverse functions to nest the data and then map a function to each row of the nested data. 

# Define season based on survey month
cod.dat$SEASON<- ifelse(cod.dat$EST_MONTH >= 2 & cod.dat$EST_MONTH <= 5, "SPRING", "FALL")
                        
# Nested dataframe by year and season
cod.nest<- cod.dat %>%
  group_by(., EST_YEAR, SEASON) %>%
  nest(., .key = "ObsData")

# Check it out -- all the data for a given year and season is collapsed and lives in the "ObsData" column
cod.nest 

# With that set up, let's write a little function hat uses each row of the ObsData to calculate the center of gravity (using SDMTools COGravity function). 
cog_map_func<- function(df){
  df.temp<- df %>%
    drop_na(DECDEG_BEGLON, DECDEG_BEGLAT, BIOMASS.MOD)
  cog.temp<- COGravity(x = df.temp$DECDEG_BEGLON, y = df.temp$DECDEG_BEGLAT, z = df.temp$BIOMASS.MOD)
  cog.out<- data.frame("Long" = cog.temp[1], "Lat" = cog.temp[3])
  return(cog.out)
}

# Now, map this function to each of the ObsData rows, creating a new COG output column
cod.nest<- cod.nest %>%
  mutate(., "COG" = map(ObsData, cog_map_func))

# Alright, now if we have a dataset with EST_YEAR, SEASON, COGx and COGy, we might be able to look at how COG has changed over time...
cod.plot.dat<- cod.nest %>%
  dplyr::select(., EST_YEAR, SEASON, COG) %>%
  unnest()

# Plot, gradient applied to year and faceted by season. For some reason, ggplot isn't liking the geom_sf and then adding points...
nelme.poly<- fortify(as(st_simplify(nelme, dTolerance = 0.001), "Spatial"))
cod.cog.plot<- ggplot() +
  geom_polygon(data = nelme.poly, aes(x = long, y = lat, group = group), color = "black", fill = NA) +
  geom_point(data = cod.plot.dat, aes(x = Long, y = Lat, color = EST_YEAR), alpha = 0.5) +
  scale_color_viridis() +
  theme_bw() +
  coord_map() +
  facet_wrap(~SEASON)

# Save it as a jpg and print it to the Markdown file
ggsave(here("Temp Results", "ExampleCodCOGPlot.jpg"), cod.cog.plot)

# Print it to Markdown file
cod.cog.plot
```

Let's see how that all worked (or didn't...)
